---
title: "simpleRegress"
author: "Taylor Stohl"
date: "Summer 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


The details to look at in this bunch of information is the residual standard error and the adjusted R-squared. The smaller the residual standard error and the larger the adjusted R-squared, the more effective the model. This model using the variable NECK appears to have a residual standard error of approximately 5.4 while the adjusted R squared is at 0.74. In other words the error was corrected by 74%.


Lets see if more effective models can be made using other variables. 


## now predict the length of a bear who has a neck of 17, 22 and 26


```{r}
new <- data.frame(NECK=c(17,22,26))
predict(simp,new)
predict(simp,new,se.fit=TRUE)
```
## Surely there is another variable that can better predict LENGTH than NECK. Find one and compare its performance to that of neck by it value of Adjusted-Rsquared (bigger is better).


First, I will model the length of a bear based on chest size to see if it produces a more effective model.


```{r}
modchest <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(modchest)
summary.lm(modchest)
```

This model is more effective than the previous as the residual standard error is smaller and the Adjusted R-squared is larger. The smaller the residual standard error the less error present. The larger the adjusted R-squared the more errors are corrected. The residual standard error in this model, modchest, is approximately 5 and the adjusted R-squared is at 0.78. The previous model titled simp had a residual standard error of approximately 5.4 and the adjusted R-squared was 0.74.
Based on these results, this model is clearly the favorable.



Next, I will use the variable HEADWTH to create a model. Let's see how effective this one will be. 




```{r}
modheadw <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(modheadw)
summary.lm(modheadw)
```


This model, modheadw, appears to be completely ineffective. The residual standard error in this model is 7.32, which is considerably large in comparison to both of the previous models simp and modchest. These numbers were 5 and 5.4. The adjusted R-squared in this model is only 0.53, whereas the first to models were 0.74 and 0.78. 



