---
title: "simpleRegress"
author: "Taylor Stohl"
date: "Summer 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```


The details to look at in this bunch of information is the residual standard error and the adjusted R-squared. The smaller the residual standard error and the larger the adjusted R-squared, the more effective the model. This model using the variable NECK appears to have a residual standard error of approximately 5.4 while the adjusted R squared is at 0.74. In other words the error was corrected by 74%.


Lets see if more effective models can be made using other variables. 


## now predict the length of a bear who has a neck of 17, 22 and 26


```{r}
new <- data.frame(NECK=c(17,22,26))
predict(simp,new)
predict(simp,new,se.fit=TRUE)
```
## Surely there is another variable that can better predict LENGTH than NECK. Find one and compare its performance to that of neck by it value of Adjusted-Rsquared (bigger is better).


First, I will model the length of a bear based on chest size to see if it produces a more effective model.


```{r}
modchest <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(modchest)
summary.lm(modchest)
```

This model is more effective than the previous as the residual standard error is smaller and the Adjusted R-squared is larger. The smaller the residual standard error the less error present. The larger the adjusted R-squared the more errors are corrected. The residual standard error in this model, modchest, is approximately 5 and the adjusted R-squared is at 0.78. The previous model titled simp had a residual standard error of approximately 5.4 and the adjusted R-squared was 0.74.
Based on these results, this model is clearly the favorable.



